---
title: "Trade_off"
author: "Khanh Duong"
date: "2024-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

**Please set working directory before running these codes.**

```{r data}
# Load necessary libraries
library(readxl)
library(dplyr)

# Load and prepare datasets
swiid9_7_summary <- read_excel("swiid9_7_summary.xlsx") %>%
  select(Country_code = code, year, gini_disp, gini_mkt) %>%
  mutate(redistribution = gini_mkt - gini_disp) %>%
  mutate(year = as.character(year))

pwt1001 <- read_excel("pwt1001.xlsx", sheet = "Data") %>%
  select(Country_code = countrycode, year, human_capital = hc, technology = ctfp) %>%
  mutate(year = as.character(year))

world_bank <- read_excel("world_bank.xlsx") %>%
  select(
    Country_code = `Country Code`,
    year = `Time Code`,
    GDP_per_capita = `NY.GDP.PCAP.KD`,
    trade = `NE.TRD.GNFS.ZS`,
    FDI = `BX.KLT.DINV.WD.GD.ZS`,
    credit = `FS.AST.PRVT.GD.ZS`,
    fertility = `SP.DYN.TFRT.IN`,
    unemployment = `SL.UEM.TOTL.ZS`,
    inflation = `FP.CPI.TOTL.ZG`
  ) %>%
  mutate(year = gsub("YR", "", as.character(year)))

# Merge datasets by 'Country_code' and 'year'
data <- pwt1001 %>%
  left_join(world_bank, by = c("Country_code", "year")) %>%
  left_join(swiid9_7_summary, by = c("Country_code", "year"))

# Load and merge country metadata, filter out rows with missing values in key columns
data <- data %>%
  filter(!is.na(GDP_per_capita), !is.na(gini_disp)) %>%
  rename(inequality = gini_disp) %>%
  select(-Country_code, -year)

min_growth <- min(data$GDP_per_capita)
max_growth <- max(data$GDP_per_capita)

min_inequality <- min(data$inequality)
max_inequality <- max(data$inequality)

```

Load and Prepare the Data

```{r}
# Load necessary libraries
library(dplyr)
library(caret)
library(glmnet)
library(mco)

# Define the predictors for the models

pre_process <- preProcess(data, method = c('bagImpute', 'range'))
data <- predict(pre_process, newdata = data)

predictors <- data %>%
  select(redistribution, human_capital, technology, trade, FDI, credit, fertility, unemployment, inflation) %>%
  as.matrix(.)

# Convert the columns to numeric, if needed
GDP_per_capita <- as.numeric(data$GDP_per_capita)
inequality <- as.numeric(data$inequality)
inequality_penalty <- data$inequality[which.max(data$GDP_per_capita)]

```

Train the Machine Learning Models (Elastic Net)

```{r message=TRUE}
# Set up cross-validation
control <- trainControl(
  method = "cv", 
  number = 5
)

tuneGrid = expand.grid(alpha = seq(0, 1, by = 0.1), lambda = 10^seq(-3, -2, length.out = 50))

# Train Elastic Net model for GDP per capita (growth)
model_growth <- train(
  predictors, GDP_per_capita,
  method = "glmnet",
  trControl = control,
  tuneGrid = tuneGrid,
  metric = "RMSE",
  na.action = na.omit
)

# Train Elastic Net model for inequality (Gini index)
model_inequality <- train(
  predictors, inequality,
  method = "glmnet",
  trControl = control,
  tuneGrid = tuneGrid,
  metric = "RMSE",
  na.action = na.omit
)

# Extract coefficients from the models
coefficients_growth <- coef(model_growth$finalModel, model_growth$bestTune$lambda)
coefficients_inequality <- coef(model_inequality$finalModel, model_inequality$bestTune$lambda)

```

Define the Optimisation Function

```{r}

objective_function <- function(x) {
  # Check for NA values in input
  if (any(is.na(x))) stop("Input vector 'x' contains NA values!")
  
  # Replace NA coefficients with zero
  growth_coeffs <- replace(coefficients_growth, is.na(coefficients_growth), 0)
  inequality_coeffs <- replace(coefficients_inequality, is.na(coefficients_inequality), 0)
  
  # Define predictors from the vector `x`
  predictors <- x[1:9]

  # Growth objective (maximize GDP per capita)
  growth <- sum(growth_coeffs * c(1, predictors))  # 1 for intercept, then the predictors
  
  # Inequality objective (minimize Gini index)
  inequality <- sum(inequality_coeffs * c(1, predictors))  # 1 for intercept, then the predictor

  # Return the objectives (negate growth for maximisation)
  return(c(-growth, inequality))
}

```

Run the Multi-Objective Optimisation Using NSGA-II

```{r}
result <- nsga2(
  objective_function,
  idim = ncol(predictors),  # 9 predictors
  odim = 2,  # Two objectives: growth and inequality
  lower.bounds = rep(0, ncol(predictors)),
  upper.bounds = rep(1, ncol(predictors))
)

# Print the Pareto front (optimal trade-offs)
pareto_front <- as.data.frame(result$value)
colnames(pareto_front) <- c("GDP_per_capita", "inequality")
pareto_front$GDP_per_capita <- -pareto_front$GDP_per_capita

```

```{r}
ggplot() +
  # Scatter plot for pareto_front data
  geom_point(data = pareto_front, aes(x = GDP_per_capita, y = inequality)) +
  xlim(0, 1) + 
  # Add smooth line
  geom_smooth(data = pareto_front, aes(x = GDP_per_capita, y = inequality), method = "loess") +
  
  geom_smooth(data = pareto_front, aes(x = GDP_per_capita, y = inequality + inequality_penalty), method = "loess") +
  
  # Scatter plot for data
  geom_point(data = data, aes(x = GDP_per_capita, y = inequality), color = "grey") + theme_minimal() +
labs(y = "Inequality", x = "Growth")

```

```{r}
# Revert the normalized values to original scale
revert <- function(x, min_val, max_val) {
  return(x * (max_val - min_val) + min_val)
}

# Print the reverted values
cat("Reverted Maximum GDP: ", revert(max(pareto_front$GDP_per_capita), min_growth, max_growth), "\n")
cat("Reverted Minimum GDP: ", revert(min(pareto_front$GDP_per_capita), min_growth, max_growth), "\n")
cat("Reverted Maximum Inequality: ", revert(max(pareto_front$inequality), min_inequality, max_inequality), "\n")
cat("Reverted Minimum Inequality: ", revert(min(pareto_front$inequality), min_inequality, max_inequality), "\n")


```

```{r}
print(cor(pareto_front$GDP_per_capita, pareto_front$inequality, method = "pearson"))

model <- lm(inequality ~ GDP_per_capita, data = pareto_front)
# Display the summary of the regression model
summary(model)

```

The gap is attributed by non-economic factors.

From an economic perspective, growth is positively associated with inequality, yet inequality levels remain within moderate bounds.
